---
title: "2. Correlation"
output: html_notebook
---

<!-- TOC -->
- [2. Correlation](#2-correlation)
  - [2.1 Purpose](#21-purpose)
  - [2.2 Expectation and the ensemble](#22-expectation-and-the-ensemble)
    - [2.2.1 Expected value](#221-expected-value)
    - [2.2.2 The ensemble and stationarity](#222-the-ensemble-and-stationarity)
    - [2.2.3 Ergodic series](#223-ergodic-series)
    - [2.2.4 Variance function](#224-variance-function)
    - [2.2.5 Autocorrelation](#225-autocorrelation)
  - [2.3 The correlogram](#23-the-correlogram)
    - [2.3.1 General discussion](#231-general-discussion)
    - [2.3.2 Example based on air passenger series](#232-example-based-on-air-passenger-series)
    - [2.3.3 Example based on the Font Reservoir series](#233-example-based-on-the-font-reservoir-series)
  - [2.4 Covariance of sums of random variables](#24-covariance-of-sums-of-random-variables)
  - [2.5 Summary of commands used in examples](#25-summary-of-commands-used-in-examples)
  - [2.6 Exercises](#26-exercises)
<!-- /TOC -->

# 2. Correlation
## 2.1 Purpose

## 2.2 Expectation and the ensemble

### 2.2.1 Expected value
```{r}
www <- "Ch02/Herald.dat"
Herald.dat <- read.table(www, header = T)
attach (Herald.dat)
```
```{r}
x <- CO; y <- Benzoa; n <- length(x)
sum((x - mean(x))*(y - mean(y))) / (n - 1)
```
```{r}
mean((x - mean(x)) * (y - mean(y)))
```
```{r}
cov(x, y)
```
```{r}
cov(x,y) / (sd(x)*sd(y))
```
```{r}
cor(x,y)
```

### 2.2.2 The ensemble and stationarity

### 2.2.3 Ergodic series

### 2.2.4 Variance function

### 2.2.5 Autocorrelation

## 2.3 The correlogram

### 2.3.1 General discussion

### 2.3.2 Example based on air passenger series
```{r}
data(AirPassengers)
AP <- AirPassengers
AP.decom <- decompose(AP, "multiplicative")
plot(ts(AP.decom$random[7:138]))
acf(AP.decom$random[7:138])
```

```{r}
sd(AP[7:138])
```
```{r}
sd(AP[7:138] - AP.decom$trend[7:138])
```
```{r}
sd(AP.decom$random[7:138])
```

### 2.3.3 Example based on the Font Reservoir series
```{r}
www <- "Ch02/Fontdsdt.dat"
Fontdsdt.dat <- read.table(www, header=T)
attach(Fontdsdt.dat)
plot(ts(adflow), ylab = ' adflow ' )
acf(adflow, xlab = ' lag (months) ' , main="")
```

## 2.4 Covariance of sums of random variables

## 2.5 Summary of commands used in examples
| mean | returns the mean (average)                                                       |
|------|----------------------------------------------------------------------------------|
| var  | returns the variance with denominator n - 1                                      |
| sd   | returns the standard deviation                                                   |
| cov  | returns the covariance with denominator n - 1                                    |
| cor  | returns the correlation                                                          |
| acf  | returns the correlogram (or sets the argument to obtain autocovariance function) |

## 2.6 Exercises
