---
title: "3. Forecasting Strategies"
output: html_notebook
---

<!-- TOC -->

- [3. Forecasting Strategies](#3-forecasting-strategies)
  - [3.1 Purpose](#31-purpose)
  - [3.2 Leading variables and associated variables](#32-leading-variables-and-associated-variables)
    - [3.2.1 Marine coatings](#321-marine-coatings)
    - [3.2.2 Building approvals publication](#322-building-approvals-publication)
    - [3.2.3 Gas supply](#323-gas-supply)
  - [3.3 Bass model](#33-bass-model)
    - [3.3.1 Background](#331-background)
    - [3.3.2 Model definition](#332-model-definition)
    - [3.3.3 Interpretation of the Bass model](#333-interpretation-of-the-bass-model)
    - [3.3.4 Example](#334-example)
  - [3.4 Exponential smoothing and the Holt-Winters method](#34-exponential-smoothing-and-the-holt-winters-method)
    - [3.4.1 Exponential smoothing](#341-exponential-smoothing)
    - [3.4.2 Holt-Winters method](#342-holt-winters-method)
    - [3.4.3 Four-year-ahead forecasts for the air passenger data](#343-four-year-ahead-forecasts-for-the-air-passenger-data)
  - [3.5 Summary of commands used in examples](#35-summary-of-commands-used-in-examples)
  - [3.6 Exercises](#36-exercises)

<!-- /TOC -->

# 3. Forecasting Strategies

## 3.1 Purpose

## 3.2 Leading variables and associated variables

### 3.2.1 Marine coatings

### 3.2.2 Building approvals publication
```{r}
www <- "Ch03/ApprovActiv.dat"
Build.dat <- read.table(www, header=T) ; attach(Build.dat)
App.ts <- ts(Approvals, start = c(1996,1), freq=4)
Act.ts <- ts(Activity, start = c(1996,1), freq=4)
ts.plot(App.ts, Act.ts, lty = c(1,3))
```
```{r}
acf(ts.union(App.ts, Act.ts))
```
```{r}
app.ran <- decompose(App.ts)$random
app.ran.ts <- window (app.ran, start = c(1996, 3) )
act.ran <- decompose (Act.ts)$random
act.ran.ts <- window (act.ran, start = c(1996, 3) )

ts.union(app.ran.ts, act.ran.ts)
class(ts.union(app.ran.ts, act.ran.ts))
length(ts.union(app.ran.ts, act.ran.ts))

acf(ts.union(app.ran.ts, act.ran.ts)[c(1:30),])
ccf(app.ran.ts, act.ran.ts)

acf()
```

```{r}
print(acf(ts.union(app.ran.ts, act.ran.ts)))
```
### 3.2.3 Gas supply

## 3.3 Bass model

### 3.3.1 Background

### 3.3.2 Model definition

### 3.3.3 Interpretation of the Bass model

### 3.3.4 Example
```{r}
T79 <- 1:10
Tdelt <- (1:100) / 10
Sales <- c(840,1470,2110,4000, 7590, 10950, 10530, 9470, 7790, 5890)
Cusales <- cumsum(Sales)
Bass.nls <- nls(Sales ~ M * ( ((P+Q)^2 / P) * exp(-(P+Q) * T79) ) / 
                  (1+(Q/P)*exp(-(P+Q)*T79))^2, start = list(M=60630, P=0.03, Q=0.38))
summary(Bass.nls)
```
```{r}
Bcoef <- coef(Bass.nls)
m <- Bcoef[1]
p <- Bcoef[2]
q <- Bcoef[3]
ngete <- exp(-(p+q) * Tdelt)
Bpdf <- m * ( (p+q)^2 / p ) * ngete / (1 + (q/p) * ngete)^2
plot(Tdelt, Bpdf, xlab = "Year from 1979",
     ylab = "Sales per year", type= 'l' )
points(T79, Sales)
Bcdf <- m * (1 - ngete)/(1 + (q/p)*ngete)
plot(Tdelt, Bcdf, xlab = "Year from 1979", ylab = "Cumulative sales", type= 'l' )
points(T79, Cusales)
```

## 3.4 Exponential smoothing and the Holt-Winters method

### 3.4.1 Exponential smoothing

* **Complaints to a motoring organisation**
```{r}
www <- "Ch03/motororg.dat"
complaints
Motor.dat <- read.table(www, header = T); attach(Motor.dat)
Comp.ts <- ts(complaints, start = c(1996, 1), freq = 12)
Comp.ts
plot(Comp.ts, xlab = "Time / months", ylab = "Complaints")
```
```{r}
Comp.hw1 <- HoltWinters(complaints, beta = 0, gamma = 0);
Comp.hw1 <- HoltWinters(Comp.ts, beta = 0, gamma = 0); 
Comp.hw1
```
```{r}
plot(Comp.hw1)
```
```{r}
Comp.hw1$SSE
```
```{r}
Comp.hw2 <- HoltWinters(complaints, alpha = 0.2, beta=0, gamma=0)
Comp.hw2 <- HoltWinters(Comp.ts, alpha = 0.2, beta=0, gamma=0)
Comp.hw2
```

```{r}
Comp.hw2$SSE
```

### 3.4.2 Holt-Winters method
* **Sales of Australian wine**
```{r}
www <- "Ch03/wine.dat"
wine.dat <- read.table(www, header = T) ; attach (wine.dat)
sweetw.ts <- ts(sweetw, start = c(1980,1), freq = 12)
plot(sweetw.ts, xlab= "Time (months)", ylab = "sales (1000 litres)")
sweetw.hw <- HoltWinters (sweetw.ts, seasonal = "mult")
sweetw.hw ; sweetw.hw$coef ; sweetw.hw$SSE
```

```{r}
sqrt(sweetw.hw$SSE/length(sweetw))
```
```{r}
sd(sweetw)
```
```{r}
plot(sweetw.hw$fitted)
plot(sweetw.hw)
```


### 3.4.3 Four-year-ahead forecasts for the air passenger data

## 3.5 Summary of commands used in examples
| nls         | non-linear least squares fit                                                |
|-------------|-----------------------------------------------------------------------------|
| HoltWinters | estimates the parameters of the Holt-Winters or exponential smoothing model |
| predict     | forecasts future values                                                     |
| ts.union    | create the union of two series                                              |
| coef        | extracts the coefficients of a fitted model                                 |

## 3.6 Exercises
