---
title: "4. Basic Stochastic Models"
output: html_notebook
---

# 4. Basic Stochastic Models

## 4.1 Purpose

## 4.2 White noise

### 4.2.1 Introduction

### 4.2.2 Definition

### 4.2.3 Simulation in R
```{r}
set.seed(1)
w <- rnorm(100)
plot(w, type = "l")
```
```{r}
x <- seq(-3,3, length = 1000)
hist(rnorm(100), prob = T); points(x, dnorm(x), type = "l")
```


### 4.2.4 Second-order properties and the correlogram

```{r}
set.seed(2)
acf(rnorm(100))
```
### 4.2.5 Fitting a white noise model

## 4.3 Random walks

### 4.3.1 Introduction

### 4.3.2 Definition

### 4.3.3 The backward shift operator

### 4.3.4 Random walk: Second-order properties

### 4.3.5 Derivation of second-order properties

### 4.3.6 The difference operator

### 4.3.7 Simulation
```{r}
x <- w <- rnorm(1000)
for (t in 2:1000) x[t] <- x[t - 1] + w[t]
plot(x, type = "l")
```

## 4.4 Fitted models and diagnostic plots

### 4.4.1 Simulated random walk series
```{r}
acf(diff(x))
```

### 4.4.2 Exchange rate series
```{r}
Z.hw <- HoltWinters(Z.ts, alpha = 1, gamma = 0)
acf(resid(Z.hw))
```

### 4.4.3 Random walk with drift
```{r}
www <- "Ch04/HP.txt"
HP.dat <- read.table(www, header = T) ; attach(HP.dat)
plot (as.ts(Price))
DP <- diff(Price) ; plot (as.ts(DP)) ; acf (DP)
mean(DP) + c(-2, 2) * sd(DP)/sqrt(length(DP))
```

## 4.5 Autoregressive models

### 4.5.1 Definition

### 4.5.2 Stationary and non-stationary AR processes

### 4.5.3 Second-order properties of an AR(1) model

### 4.5.4 Derivation of second-order properties for an AR(1) process

### 4.5.5 Correlogram of an AR(1) process
```{r}
rho <- function(k, alpha) alpha^k
# layout(1:1)
plot(0:10, rho(0:10, 0.7), type = "b")
plot(0:10, rho(0:10, -0.7), type = "b")
```

### 4.5.6 Partial autocorrelation

### 4.5.7 Simulation
```{r}
set.seed(1)
x <- w <- rnorm(100)
for (t in 2:100) x[t] <- 0.7 * x[t - 1] + w[t]
plot(x, type = "l")
acf(x)
pacf(x)
```

## 4.6 Fitted models

### 4.6.1 Model fitted to simulated series
```{r}
x.ar <- ar(x, method = "mle")
x.ar$order
```
```{r}
x.ar$ar + c(-2, 2) * sqrt(x.ar$asy.var)
```

### 4.6.2 Exchange rate series: Fitted AR model
```{r}
Z.ar <- ar(Z.ts)
mean(Z.ts)
Z.ar$order
Z.ar$ar
Z.ar$ar + c(-2, 2) * sqrt(Z.ar$asy.var)
acf(Z.ar$res[-1])
```
### 4.6.3 Global temperature series: Fitted AR model
```{r}
www = "Ch04/global.dat"
Global = scan(www)
Global.ts = ts(Global, st = c(1856, 1), end = c(2005, 12), fr = 12)
```
```{r}
Global.ar <- ar(aggregate(Global.ts, FUN = mean), method = "mle")
mean(aggregate(Global.ts, FUN = mean))
Global.ar$order
Global.ar$ar
acf(Global.ar$res[-(1:Global.ar$order)], lag = 50)
```


## 4.7 Summary of R commands

| set.seed | sets a seed for the random number generator enabling a simulation to be reproduced |
|----------|------------------------------------------------------------------------------------|
| rnorm    | simulates Gaussian white noise series                                              |
| diff     | creates a series of first-order differences                                        |
| ar       | gets the best fitting AR( p) model                                                 |
| pacf     | extracts partial autocorrelations and partial correlogram                          |
| polyroot | extracts the roots of a polynomial                                                 |
| resid    | extracts the residuals from a fitted model                                         |

## 4.8 Exercises

