---
title: "6. Stationary Models"
output: html_notebook
---
<!-- TOC -->

- [6. Stationary Models](#6-stationary-models)
  - [6.1 Purpose](#61-purpose)
  - [6.2 Strictly stationary series](#62-strictly-stationary-series)
  - [6.3 Moving average models](#63-moving-average-models)
    - [6.3.1 MA(q) process: Definition and properties](#631-maq-process-definition-and-properties)
    - [6.3.2 R examples: Correlogram and simulation](#632-r-examples-correlogram-and-simulation)
  - [6.4 Fitted MA models](#64-fitted-ma-models)
    - [6.4.1 Model fitted to simulated series](#641-model-fitted-to-simulated-series)
    - [6.4.2 Exchange rate series: Fitted MA model](#642-exchange-rate-series-fitted-ma-model)
  - [6.5 Mixed models: The ARMA process](#65-mixed-models-the-arma-process)
    - [6.5.1 Definition](#651-definition)
    - [6.5.2 Derivation of second-order properties](#652-derivation-of-second-order-properties)
  - [6.6 ARMA models: Empirical analysis](#66-arma-models-empirical-analysis)
    - [6.6.1 Simulation and fitting](#661-simulation-and-fitting)
    - [6.6.2 Exchange rate series](#662-exchange-rate-series)
    - [6.6.3 Electricity production series](#663-electricity-production-series)
    - [6.6.4 Wave tank data](#664-wave-tank-data)
  - [6.7 Summary of R commands](#67-summary-of-r-commands)
  - [6.8 Exercises](#68-exercises)

<!-- /TOC -->

# 6. Stationary Models
## 6.1 Purpose

## 6.2 Strictly stationary series

## 6.3 Moving average models

### 6.3.1 MA(q) process: Definition and properties

### 6.3.2 R examples: Correlogram and simulation
```{r}
rho <- function(k, beta) {
q <- length(beta) - 1
if (k > q) ACF <- 0 else {
  s1 <- 0; s2 <- 0
  for (i in 1:(q-k+1)) s1 <- s1 + beta[i] * beta[i+k]
  for (i in 1:(q+1)) s2 <- s2 + beta[i]^2
  ACF <- s1 / s2}
ACF}
```

```{r}
beta <- c(1, 0.7, 0.5, 0.2)
rho.k <- rep(1, 10)
for (k in 1:10) rho.k[k] <- rho(k, beta)
plot(0:10, c(1, rho.k), pch = 4, ylab = expression(rho[k]))
abline(0, 0)
```


```{r}
set.seed(1)
b <- c(0.8, 0.6, 0.4)
x <- w <- rnorm(1000)
for (t in 4:1000) {
for (j in 1:3) x[t] <- x[t] + b[j] * w[t - j]}
plot(x, type = "l")
acf(x)
```

## 6.4 Fitted MA models

### 6.4.1 Model fitted to simulated series
```{r}
x.ma <- arima(x, order = c(0, 0, 3))
x.ma
```



### 6.4.2 Exchange rate series: Fitted MA model
```{r}
www <- "Ch06/pounds_nz.dat"
x <- read.table(www, header = T)
x.ts <- ts(x, st = 1991, fr = 4)
x.ma <- arima(x.ts, order = c(0, 0, 1))
x.ma
```

```{r}
acf(x.ma$res[-1])
```


## 6.5 Mixed models: The ARMA process

### 6.5.1 Definition

### 6.5.2 Derivation of second-order properties

## 6.6 ARMA models: Empirical analysis

### 6.6.1 Simulation and fitting

```{r}
set.seed(1)
x <- arima.sim(n = 10000, list(ar = -0.6, ma = 0.5))
coef(arima(x, order = c(1, 0, 1)))
```
### 6.6.2 Exchange rate series

```{r}
x.ma <- arima(x.ts, order = c(0, 0, 1))
x.ar <- arima(x.ts, order = c(1, 0, 0))
x.arma <- arima(x.ts, order = c(1, 0, 1))
AIC(x.ma)
AIC(x.ar)
```

```{r}
AIC(x.arma)
x.arma
acf(resid(x.arma))
```

### 6.6.3 Electricity production series
```{r}
www <- "Ch06/cbe.dat"
CBE <- read.table(www, header = T)
Elec.ts <- ts(CBE[, 3], start = 1958, freq = 12)
Time <- 1:length(Elec.ts)
Imth <- cycle(Elec.ts)
Elec.lm <- lm(log(Elec.ts) ~ Time + I(Time^2) + factor(Imth))
acf(resid(Elec.lm))
```


```{r}
best.order <- c(0, 0, 0)
best.aic <- Inf

for (i in 0:2) for (j in 0:2) {
  fit.aic <- AIC(arima(resid(Elec.lm), order = c(i, 0, j)))
  if (fit.aic < best.aic) {
    best.order <- c(i, 0, j)
    best.arma <- arima(resid(Elec.lm), order = best.order)
    best.aic <- fit.aic
  }
}
```

```{r}
best.order
```
```{r}
acf(resid(best.arma))
```
```{r}
new.time <- seq(length(Elec.ts), length = 36)
new.data <- data.frame(Time = new.time, Imth = rep(1:12,3))
predict.lm <- predict(Elec.lm, new.data)
predict.arma <- predict(best.arma, n.ahead = 36)
elec.pred <- ts(exp(predict.lm + predict.arma$pred), start = 1991, freq = 12)
ts.plot(cbind(Elec.ts, elec.pred), lty = 1:2)
```

### 6.6.4 Wave tank data
```{r}
www <- "Ch06/wave.dat"
wave.dat <- read.table(www, header = T)
attach (wave.dat)
layout(1:3)
plot (as.ts(waveht), ylab = ' Wave height ' )
acf (waveht)
pacf (waveht)
wave.arma <- arima(waveht, order = c(4,0,4))
acf (wave.arma$res[-(1:4)])
pacf (wave.arma$res[-(1:4)])
hist(wave.arma$res[-(1:4)], xlab= ' height / mm ' , main= '' )
```


## 6.7 Summary of R commands
| arima.sim  | simulates data from an ARMA (or ARIMA) process |
|------------|------------------------------------------------|
| arima      | fits an ARMA (or ARIMA) model to data          |
| seq        | generates a sequence                           |
| expression | used to plot maths symbol                      |

## 6.8 Exercises
