---
title: "7. Non-stationary Models"
output: html_notebook
--- 

<!-- TOC -->

- [7. Non-stationary Models](#7-non-stationary-models)
  - [7.1 Purpose](#71-purpose)
  - [7.2 Non-seasonal ARIMA models](#72-non-seasonal-arima-models)
    - [7.2.1 Differencing and the electricity series](#721-differencing-and-the-electricity-series)
    - [7.2.2 Integrated model](#722-integrated-model)
    - [7.2.3 Definition and examples](#723-definition-and-examples)
    - [7.2.4 Simulation and fitting](#724-simulation-and-fitting)
    - [7.2.5 IMA(1, 1) model fitted to the beer production series](#725-ima1-1-model-fitted-to-the-beer-production-series)
  - [7.3 Seasonal ARIMA models](#73-seasonal-arima-models)
    - [7.3.1 Definition](#731-definition)
    - [7.3.2 Fitting procedure](#732-fitting-procedure)
  - [7.4 ARCH models](#74-arch-models)
    - [7.4.1 S&P500 series](#741-sp500-series)
    - [7.4.2 Modelling volatility: Definition of the ARCH model](#742-modelling-volatility-definition-of-the-arch-model)
    - [7.4.3 Extensions and GARCH models](#743-extensions-and-garch-models)
    - [7.4.4 Simulation and fitted GARCH model](#744-simulation-and-fitted-garch-model)
    - [7.4.5 Fit to S&P500 series](#745-fit-to-sp500-series)
    - [7.4.6 Volatility in climate series](#746-volatility-in-climate-series)
    - [7.4.7 GARCH in forecasts and simulations](#747-garch-in-forecasts-and-simulations)
  - [7.5 Summary of R commands](#75-summary-of-r-commands)
  - [7.6 Exercises](#76-exercises)

<!-- /TOC -->
# 7. Non-stationary Models
## 7.1 Purpose

## 7.2 Non-seasonal ARIMA models

###7.2.1 Differencing and the electricity series
```{r}
www <- "Ch07/cbe.dat"
CBE <- read.table(www, he = T)
Elec.ts <- ts(CBE[, 3], start = 1958, freq = 12)
layout(c(1, 1, 2, 3))
plot(Elec.ts)
plot(diff(Elec.ts))
plot(diff(log(Elec.ts)))
```

###7.2.2 Integrated model

###7.2.3 Definition and examples

###7.2.4 Simulation and fitting
```{r}
set.seed(1)
x <- w <- rnorm(1000)
for (i in 3:1000) 
  x[i] <- 0.5 * x[i - 1] + x[i - 1] - 0.5 * x[i - 2] + w[i] + 0.3 * w[i - 1]
arima(x, order = c(1, 1, 1))
```

```{r}      
x <- arima.sim(model = list(order = c(1, 1, 1), ar = 0.5, ma = 0.3), n = 1000)
arima(x, order = c(1, 1, 1))
```


###7.2.5 IMA(1, 1) model fitted to the beer production series
```{r}      
Beer.ts <- ts(CBE[, 2], start = 1958, freq = 12)
Beer.ima <- arima(Beer.ts, order = c(0, 1, 1))
Beer.ima
```

```{r}
acf(resid(Beer.ima))
```


```{r}
Beer.1991 <- predict(Beer.ima, n.ahead = 12)
sum(Beer.1991$pred)
```
## 7.3 Seasonal ARIMA models

### 7.3.1 Definition

### 7.3.2 Fitting procedure
```{r}
AIC (arima(log(Elec.ts), order = c(1,1,0), 
           seas = list(order = c(1,0,0), 12)))

AIC (arima(log(Elec.ts), order = c(0,1,1),
           seas = list(order = c(0,0,1), 12)))
```

```{r}
get.best.arima <- function(x.ts, maxord = c(1,1,1,1,1,1))
{
  best.aic <- 1e8
  n <- length(x.ts)
  for (p in 0:maxord[1]) for(d in 0:maxord[2]) for(q in 0:maxord[3])
    for (P in 0:maxord[4]) for(D in 0:maxord[5]) for(Q in 0:maxord[6])
    {
      fit <- arima(x.ts, 
                   order = c(p,d,q),
                   seas = list(order = c(P,D,Q),
                               frequency(x.ts)), method = "CSS")
      fit.aic <- -2 * fit$loglik + (log(n) + 1) * length(fit$coef)
      if (fit.aic < best.aic)
        {
          best.aic <- fit.aic
          best.fit <- fit
          best.model <- c(p,d,q,P,D,Q)
        }
    }
  list(best.aic, best.fit, best.model)
}
  
best.arima.elec <- get.best.arima(log(Elec.ts), maxord = c(2,2,2,2,2,2))
best.fit.elec <- best.arima.elec[[2]]
acf(resid(best.fit.elec))
best.arima.elec [[3]]
ts.plot( cbind( window(Elec.ts,start = 1981),
                exp(predict(best.fit.elec,12)$pred) ), lty = 1:2)
```  

## 7.4 ARCH models

### 7.4.1 S&P500 series
```{r}
library(MASS)
data(SP500)
plot(SP500, type = 'l' )
acf(SP500)
```

```{r}
acf((SP500 - mean(SP500))^2)
```

### 7.4.2 Modelling volatility: Definition of the ARCH model

### 7.4.3 Extensions and GARCH models

### 7.4.4 Simulation and fitted GARCH model

```{r}
set.seed(1)
alpha0 <- 0.1
alpha1 <- 0.4
beta1 <- 0.2
w <- rnorm(10000)
a <- rep(0, 10000)
h <- rep(0, 10000)
for (i in 2:10000) {
  h[i] <- alpha0 + alpha1 * (a[i - 1]^2) + beta1 * h[i - 1]
  a[i] <- w[i] * sqrt(h[i])
}
acf(a)
acf(a^2)
```

```{r}
library(tseries)
a.garch <- garch(a, grad = "numerical", trace = FALSE)
confint(a.garch)
```
### 7.4.5 Fit to S&P500 series
```{r}
sp.garch <- garch(SP500, trace = F)
sp.res <- sp.garch$res[-1]
acf(sp.res)
acf(sp.res^2)
```

### 7.4.6 Volatility in climate series
```{r}
stemp <- scan("Ch07/stemp.dat")
stemp.ts <- ts(stemp, start = 1850, freq = 12)
plot(stemp.ts)
stemp.best <- get.best.arima(stemp.ts, maxord = rep(2,6))
stemp.best[[3]]

stemp.arima <- arima(stemp.ts, order = c(1,1,2),
seas = list(order = c(2,0,1), 12))
```

```{r}
t( confint(stemp.arima) )
```

```{r}
stemp.arima <- arima(stemp.ts, order = c(1,1,2),
                     seas = list(order = c(1,0,1), 12))
t( confint(stemp.arima) )
```

```{r}
stemp.res <- resid(stemp.arima)
layout(1:2)
```
```{r}
acf(stemp.res)
acf(stemp.res^2)
```

```{r}
stemp.garch <- garch(stemp.res, trace = F)
t(confint(stemp.garch))
```

```{r}
stemp.garch.res <- resid(stemp.garch)[-1]
acf(stemp.garch.res)
acf(stemp.garch.res^2)
```

### 7.4.7 GARCH in forecasts and simulations

## 7.5 Summary of R commands
* garch fits a GARCH (or ARCH) model to data

## 7.6 Exercises
