---
title: "9. Spectral Analysis"
output: html_notebook
--- 
<!-- TOC -->

- [9. Spectral Analysis](#9-spectral-analysis)
  - [9.1 Purpose](#91-purpose)
  - [9.2 Periodic signals](#92-periodic-signals)
    - [9.2.1 Sine waves](#921-sine-waves)
    - [9.2.2 Unit of measurement of frequency](#922-unit-of-measurement-of-frequency)
  - [9.3 Spectrum](#93-spectrum)
    - [9.3.1 Fitting sine waves](#931-fitting-sine-waves)
    - [9.3.2 Sample spectrum](#932-sample-spectrum)
  - [9.4 Spectra of simulated series](#94-spectra-of-simulated-series)
    - [9.4.1 White noise](#941-white-noise)
    - [9.4.2 AR(1): Positive coefficient](#942-ar1-positive-coefficient)
    - [9.4.3 AR(1): Negative coefficient](#943-ar1-negative-coefficient)
    - [9.4.4 AR(2)](#944-ar2)
  - [9.5 Sampling interval and record length](#95-sampling-interval-and-record-length)
    - [9.5.1 Nyquist frequency](#951-nyquist-frequency)
    - [9.5.2 Record length](#952-record-length)
  - [9.6 Applications](#96-applications)
    - [9.6.1 Wave tank data](#961-wave-tank-data)
    - [9.6.2 Fault detection on electric motors](#962-fault-detection-on-electric-motors)
    - [9.6.3 Measurement of vibration dose](#963-measurement-of-vibration-dose)
    - [9.6.4 Climatic indices](#964-climatic-indices)
    - [9.6.5 Bank loan rate](#965-bank-loan-rate)
  - [9.7 Discrete Fourier transform (DFT](#97-discrete-fourier-transform-dft)
  - [9.8 The spectrum of a random process](#98-the-spectrum-of-a-random-process)
    - [9.8.1 Discrete white noise](#981-discrete-white-noise)
    - [9.8.2 AR](#982-ar)
    - [9.8.3 Derivation of spectrum](#983-derivation-of-spectrum)
  - [9.9 Autoregressive spectrum estimation](#99-autoregressive-spectrum-estimation)
  - [9.10 Finer details](#910-finer-details)
    - [9.10.1 Leakage](#9101-leakage)
    - [9.10.2 Confidence intervals](#9102-confidence-intervals)
    - [9.10.3 Daniell windows](#9103-daniell-windows)
    - [9.10.4 Padding](#9104-padding)
    - [9.10.5 Tapering](#9105-tapering)
    - [9.10.6 Spectral analysis compared with wavelets](#9106-spectral-analysis-compared-with-wavelets)
  - [9.11 Summary of additional commands used](#911-summary-of-additional-commands-used)
  - [9.12 Exercises](#912-exercises)

<!-- /TOC -->

# 9. Spectral Analysis
## 9.1 Purpose

## 9.2 Periodic signals

### 9.2.1 Sine waves

### 9.2.2 Unit of measurement of frequency

## 9.3 Spectrum

### 9.3.1 Fitting sine waves

### 9.3.2 Sample spectrum

## 9.4 Spectra of simulated series

### 9.4.1 White noise
```{r}
layout(1:2)
set.seed(1)
x <- rnorm(2048)
spectrum(x, log = c("no"))
spectrum(x, span = 65, log = c("no"))
```


```{r}
x.spec <- spectrum (x, span = 65, log = c("no"))
spx <- x.spec$freq # Del
spy <- 2 * x.spec$spec
plot (spx, spy, xlab = "Hz", ylab = "variance/Hz", type = "l")
```

### 9.4.2 AR(1): Positive coefficient
```{r}
set.seed(1)
x <- w <- rnorm(1024)
for (t in 2:1024) x[t]<- 0.9 * x[t-1] + w[t]
layout(1:3)
plot(as.ts(x))
acf(x)
spectrum(x, span = 51, log = c("no"))
```
### 9.4.3 AR(1): Negative coefficient

### 9.4.4 AR(2)
```{r}
set.seed(1)
x <- w <- rnorm(1024)
for (t in 3:1024) x[t] <- x[t-1] - 0.6 * x[t-2] + w[t]
layout (1:3)
plot (as.ts(x))
acf (x)
spectrum (x, span = 51, log = c("no"))
```

## 9.5 Sampling interval and record length

### 9.5.1 Nyquist frequency
```{r}
t <- (0:10) / 5
tc <- (0:2000) / 1000
x <- sin (2 * pi * t)
xc <- sin (2 * pi * tc)
xa <- sin (-4 * 2 * pi * tc)
plot (t, x)
lines (tc, xc)
lines (tc, xa, lty = "dashed")
```

### 9.5.2 Record length

## 9.6 Applications

### 9.6.1 Wave tank data
```{r}
www <- "Ch09/wave.dat"
wavetank.dat <- read.table(www, header=T)
# attach (wavetank.dat)
layout (1:3)
plot (as.ts(waveht))
acf (waveht)
spectrum (waveht)
```
### 9.6.2 Fault detection on electric motors
```{r}
www <- "Ch09/imotor.txt"
imotor.dat <- read.table(www, header = T)
attach (imotor.dat)
xg.spec <- spectrum(good, span = 9)
xb.spec <- spectrum(broken, span = 9)
freqg <- 400 * xg.spec$freq [4400:5600]
freqb <- 400 * xb.spec$freq [4400:5600]
plot(freqg, 10*log10(xg.spec$spec[4400:5600]), main = "",
xlab = "Frequency (Hz)", ylab = "Current spectrum (dB)", type="l")
lines(freqb, 10 * log10(xb.spec$spec[4400:5600]), lty = "dashed")
sd(good)
sd(broken)
```

### 9.6.3 Measurement of vibration dose


```{r}
www <- "Ch09/zdd.txt"
zdotdot.dat <- read.table(www, header = T)
attach (zdotdot.dat)
www <- "Ch09/vibdoswt.txt"
wt.dat <- read.table (www, header = T)
attach (wt.dat)
acceln.spec <- spectrum (Accelnz, span = sqrt(2 * length(Accelnz)))
Frequ <- 200 * acceln.spec$freq
Sord <- 2 * acceln.spec$spec / 200
Time <- (1:1000) / 200
layout (1:3)
plot (Time, Accelnz, xlab = "Time (s)",
ylab = expression(mm~ s^-2),
main = "Acceleration", type = "l")

plot (Frequ, Sord, main = "Spectrum", xlab = "Frequency (Hz)",
      ylab = expression(mm^2~s^-4~Hz^-1), type = "l")
plot (Frequ, Weight, xlab = "Frequency (Hz)",

      main = "Weighting function", type = "l")

sd (Accelnz)
sqrt( sum(Sord * Weight) * 0.2 )
```

```{r}
demo(plotmath) 
```

### 9.6.4 Climatic indices
* More details and the data are at http://www.cru.uea.ac.uk/cru/data/soi.htm.  

```{r}
www <- "Ch09/soi.txt"
soi.dat <- read.table(www, header = T)
attach (soi.dat)
soi.ts <- ts(SOI, st = c(1866, 1), end = c(2006, 11), fr = 12)
layout (1:3)
plot (soi.ts)
soi.spec <- spectrum( SOI, span = sqrt(2 * length(SOI)) )
plot (soi.spec$freq[1:60], soi.spec$spec[1:60], type = "l")
```
* The time series data are available from http://jisao.washington.edu/pdo/.  

```{r}
www <- "Ch09/pdo.txt"
pdo.dat <- read.table(www, header = T)
attach (pdo.dat)
pdo.ts <- ts( PDO, st = c(1900, 1), end = c(2007, 11), fr = 12 )
layout (1:2)
plot (pdo.ts)
spectrum( PDO, span = sqrt( 2 * length(PDO) ) )
```

### 9.6.5 Bank loan rate
```{r}
www <- "Ch09/mprime.txt"
intr.dat <- read.table(www, header = T)
attach (intr.dat)
layout (1:3)
plot (as.ts(Interest), ylab = ' Interest rate ' )
acf (Interest)
spectrum(Interest, span = sqrt(length(Interest)) / 4)
```
## 9.7 Discrete Fourier transform (DFT
```{r}
z1 <- 2 + (0+3i)
z2 <- -1 - (0+1i)
z1 - z2
z1 * z2
abs(z1)
```


```{r}
set.seed(1)
n <- 8
x <- rnorm(n)
x
X <- fft(x)
X
fft(X, inverse = TRUE)/n
```

```{r}
fq <- 2 * pi / n
frq <- 0
FL <- 0
FL [1] <- X[1]^2 / n^2
frq[1] <- 0
for ( j in 2:(n/2) ) {
  FL [j] <- 2 * (X[j] * X[n+2-j]) / n^2
  frq[j] <- fq * (j-1)
}
FL [n/2 + 1] <- X[n/2 + 1]^2 / n^2
frq[n/2 + 1] <- pi
```

## 9.8 The spectrum of a random process

### 9.8.1 Discrete white noise

### 9.8.2 AR

### 9.8.3 Derivation of spectrum

## 9.9 Autoregressive spectrum estimation

## 9.10 Finer details

### 9.10.1 Leakage

### 9.10.2 Confidence intervals

### 9.10.3 Daniell windows

### 9.10.4 Padding

### 9.10.5 Tapering

### 9.10.6 Spectral analysis compared with wavelets

## 9.11 Summary of additional commands used
| spectrum  | returns the spectrum                                 |
|-----------|------------------------------------------------------|
| spec.pgam | returns the spectrum with more control of parameters |
| fft       | returns the DFT                                      |

## 9.12 Exercises
