---
title: "11. Multivariate Models"
output: html_notebook
--- 
<!-- TOC -->

- [11. Multivariate Models](#11-multivariate-models)
  - [11.1 Purpose](#111-purpose)
  - [11.2 Spurious regression](#112-spurious-regression)
  - [11.3 Tests for unit roots](#113-tests-for-unit-roots)
  - [11.4 Cointegration](#114-cointegration)
    - [11.4.1 Definition](#1141-definition)
    - [11.4.2 Exchange rate series](#1142-exchange-rate-series)
  - [11.5 Bivariate and multivariate white noise](#115-bivariate-and-multivariate-white-noise)
  - [11.6 Vector autoregressive models](#116-vector-autoregressive-models)
    - [11.6.1 VAR model fitted to US economic series](#1161-var-model-fitted-to-us-economic-series)
  - [11.7 Summary of R commands](#117-summary-of-r-commands)
  - [11.8 Exercises](#118-exercises)

<!-- /TOC -->
# 11. Multivariate Models

## 11.1 Purpose

## 11.2 Spurious regression
```{r}
www <- "Ch11/cbe.dat"
CBE <- read.table(www, header = T)
Elec.ts <- ts(CBE[, 3], start = 1958, freq = 12)
Choc.ts <- ts(CBE[, 1], start = 1958, freq = 12)
plot(as.vector(aggregate(Choc.ts)), as.vector(aggregate(Elec.ts)))
cor(aggregate(Choc.ts), aggregate(Elec.ts))
```

```{r}
set.seed(10); x <- rnorm(100); y <- rnorm(100)
for(i in 2:100) {
  x[i] <- x[i-1] + rnorm(1)
  y[i] <- y[i-1] + rnorm(1) 
  }
plot(x, y)
cor(x, y)
```


```{r}
www <- "Ch11/us_rates.dat"
xrates <- read.table(www, header = T)
xrates[1:3, ]
acf( diff(xrates$UK) )
acf( diff(xrates$EU) )
plot(xrates$UK, xrates$EU, pch = 4)
cor(xrates$UK, xrates$EU)
0.946
```

## 11.3 Tests for unit roots
```{r}
library(tseries)
adf.test(x)
```

```{r}
pp.test(xrates$UK)
```
```{r}
pp.test(xrates$EU)
```

## 11.4 Cointegration

### 11.4.1 Definition
```{r}
x <- y <- mu <- rep(0, 1000)
for (i in 2:1000) mu[i] <- mu[i - 1] + rnorm(1)
x <- mu + rnorm(1000)
y <- mu + rnorm(1000)
adf.test(x)$p.value
adf.test(y)$p.value
po.test(cbind(x, y))
```
### 11.4.2 Exchange rate series
```{r}
po.test(cbind(xrates$UK, xrates$EU))
ukeu.lm <- lm(xrates$UK ~ xrates$EU)
ukeu.res <- resid(ukeu.lm)
ukeu.res.ar <- ar(ukeu.res)
ukeu.res.ar$order
AIC(arima(ukeu.res, order = c(3, 0, 0)))
AIC(arima(ukeu.res, order = c(2, 0, 0)))
AIC(arima(ukeu.res, order = c(1, 0, 0)))
AIC(arima(ukeu.res, order = c(1, 1, 0)))
```

## 11.5 Bivariate and multivariate white noise
```{r}
library(mvtnorm)
cov.mat <- matrix(c(1, 0.8, 0.8, 1), nr = 2)
w <- rmvnorm(1000, sigma = cov.mat)
cov(w)
wx <- w[, 1]
wy <- w[, 2]
ccf(wx, wy, main = "")
```


## 11.6 Vector autoregressive models
```{r}
Mod(polyroot(c(1, -0.5, -0.02)))
x <- y <- rep(0, 1000)
x[1] <- wx[1]
y[1] <- wy[1]
for (i in 2:1000) {
  x[i] <- 0.4 * x[i - 1] + 0.3 * y[i - 1] + wx[i]
  y[i] <- 0.2 * x[i - 1] + 0.1 * y[i - 1] + wy[i]
}
xy.ar <- ar(cbind(x, y))
xy.ar$ar[, , ]
```

### 11.6.1 VAR model fitted to US economic series

```{r}
library(tseries)
data(USeconomic)
US.ar <- ar(cbind(GNP, M1), method="ols", dmean=T, intercept=F)
US.ar$ar
```
```{r}
acf(US.ar$res[-c(1:3), 1])
acf(US.ar$res[-c(1:3), 2])
```
```{r}
US.var <- VAR(cbind(GNP, M1), p = 2, type = "trend")
coef(US.var)
```
```{r}
acf(resid(US.var)[, 1])
acf(resid(US.var)[, 2])
```
```{r}
US.pred <- predict(US.var, n.ahead = 4)
US.pred
```

```{r}
GNP.pred <- ts(US.pred$fcst$GNP[, 1], st = 1988, fr = 4)
M1.pred <- ts(US.pred$fcst$M1[, 1], st = 1988, fr = 4)
ts.plot(cbind(window(GNP, start = 1981), GNP.pred), lty = 1:2)
ts.plot(cbind(window(M1, start = 1981), M1.pred), lty = 1:2)
```


## 11.7 Summary of R commands
| adf.test | Dickey-Fuller test for unit roots                                 |
|----------|-------------------------------------------------------------------|
| pp.test  | Phillips-Perron test for unit roots                               |
| rmvnorm  | multivariate white noise simulation                               |
| po.test  | Phillips-Ouliaris cointegration test                              |
| ar       | Fits the VAR model based on the smallest AIC                      |
| VAR      | Fits the VAR model based on least squares (vars package required) |

## 11.8 Exercises
